### 2018-12-14 report/paper.tex
report

% \documentclass[sigplan,10pt,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
% \documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
\documentclass[sigplan,10pt]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}

%% Conference information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmConference[Purdue CS590-DLS]{}{Dec 2019}{West Lafayette, Indiana}
\acmYear{2019}
\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%\citestyle{acmauthoryear}  %% For author/year citations
%\citestyle{acmnumeric}     %% For numeric citations
%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
                            %% sorting of references within a single citation;
                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
                            %% rendered as [14,5,2] rather than [2,5,14].
%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
                            %% compression of sequential references within a
                            %% single citation;
                            %% e.g., \cite{Baker12,Baker14,Baker16}
                            %% rendered as [2,3,4] rather than [2-4].

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from traditional SIGPLAN
%% proceedings format to PACMPL format must update the
%% '\documentclass' and topmatter commands above; see
%% 'acmart-pacmpl-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[scaled=0.8]{beramono}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{xcolor,colortbl}
\usepackage{url}
\usepackage{listings}
\usepackage{paralist}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{flushend}
\usepackage{bcprules}
\usepackage{nicefrac}
\usepackage{xfrac}
\input{macros}

\begin{document}

%% Title information
\title{Learning Heuristics of SAT by Graph Neural Networks}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
% \subtitle{Subtitle}                     %% \subtitle is optional
% \subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

\author
[Guannan Wei]
{Guannan Wei \\ Purdue University}

\lstMakeShortInline[keywordstyle=,%
              flexiblecolumns=false,%
              %basewidth={0.56em, 0.52em},%
              mathescape=false,%
              basicstyle=\tt]@

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  In this document, we summarize the discussion and provide some formal description of the proposed model.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\renewcommand\thefootnotecopyrightpermission{}
\footnotetextcopyrightpermission{Copyright held by the authors.}
\fancyhead[RO,LE]{Preprint, Dec 2019}

\section{Problem Statement}

In this paper, we report the progress and result on the project that aims to learn branching heuristics 
for SAT problem based on DPLL algorithm. The DPLL algorithm \cite{Davis:1962:MPT:368273.368557} and its 
variants are widely used in many modern SAT solvers. 
The core ideas behind DPLL are searching and unit propagation. 
The searching involves choosing a variable and assigning it to either $\top$ or $\bot$, 
which is the part of our concern in this paper.

The DPLL algorithm works on Conjunction Normal Form. A CNF formula is a conjunction of clauses. A clause is a disjunction of literals. A literal is either a variable (symbol) or a negation of a variable. Formulas in other forms can be translated to CNF \cite{tseitin1983complexity}. For example, $(x_1 \vee x_2) \wedge (\neg x_2 \vee x_3)$ is a valid CNF formula.

The pseudocode of DPLL procedure is listed as follows:

\begin{lstlisting}
def dpll(f: Formula): Boolean = {
  if (f.hasUnsatClause) return false
  if (f.isSat) return true
  if (f.hasUnitClause) return dpll(f.elimUnitClause)
  val v = f.pickVariable
  val tryTrue = dpll(f.addUnitClause(v))
  if (tryTrue) true
  else dpll(f.addUnitClause(neg(v)))
}
\end{lstlisting}

Particularly, we would like to learn the branching heuristics for \texttt{f.pickVariable}.
In this paper, we use graph neural network to encode a CNF formula and use reinforcement learning 
to learn the best branching heuristics.
We use the policy gradient method for the reinforcement learning part.

\section{Proposed Model}

\subsection{General Description}

\paragraph{Graph Representation}

The graph embedding captures the permutation invariance and the structure of a CNF formula.
We first review NeuroSAT's graph embedding of a CNF formula \cite{selsam2018learning}.

Given a formula $F$ of $m$ clauses and $n$ variables in CNF, 
one can construct a graph $G = \langle V, E \rangle$ such that 
$V = V_{lit} \cup V_{cls}$ where $V_{lit}$ is the set of vertices embedding literals,
$V_{cls}$ is the set of vertices embedding clauses. 
And $E$ contains two different types of edge, the first one is the edegs connecting literals and its
containing clauses, the second one is the edges connecting a variable and its negation, formally,
$ E = \{ (l, c) | l \in c \} \cup \{ (l, \neg l) | l \}$ for all literals $l$ and clauses $c$.
Note that in the NeuroSAT's encoding, there is a vertex for every variable and its negation no matter
it appears or not, so the size of vertices set $|V| = 2n + m$.

\paragraph{Message-Passing Neural Network}

After the graph is constructed, NeuroSAT uses message-passing mechanism to refine the vector space embeddings.
One iteration of a message-passing consists of two stages:
1) first every clause receives messages from its neighboring literals and update the clause embedding;
2) then every literal receives its message from its neighboring clauses and its negated literal vertices,
and update its embedding.

Suppose the size of embedding vectors is $d$. The initial embedding vectors of literals and clauses 
are $L^{0} \in \mathbb{R}^{2n \times d}$ and $C^{0} \in \mathbb{R}^{m \times d}$ which are randomly initialized.
NeuroSAT uses two MLPs $\mathcal{L}_{msg}$ and $\mathcal{C}_{msg}$ to produce the messages from embeddings, 
and two LSTMs $\mathcal{L}_u$ and $\mathcal{C}_u$ to update the embeddings every iteration 
(the hidden states are elided). 
As mentioned above, for the time $t$ there are two stages in a iteration:
$$ C^{t+1} \leftarrow \mathcal{C}_u( \mathcal{L}_{msg}(L^t) ) $$
$$ L^{t+1} \leftarrow \mathcal{L}_u( \mathcal{C}_{msg}(C^{t+1}) ) $$

After $T$ iterations, with the embedding produced, another MLP $\mathcal{P}$ which served as the policy network takes 
the embeddings as input and produces the vector $\mathcal{P}(L^T) \in \mathbb{R}^{2n}$ that represents the probabilities of choosing which literal.

\subsection{Embedding Problem}

However, naively adapting NeuroSAT's graph embedding is not going to work.
Because NeuroSAT models SAT as a supervised learning problem without using traditional algorithms 
such as DPLL, which then means the graph structure is static. 
But in the DPLL and reinforcement learning setting, the formula is changed every time when we make a decision --
some variables and clauses may completely be eliminated by unit propagation over the time.

This further raises two entangled problems:
1) How to embed the assignments as well as the decision history into the graph.
2) How to reflect the structural changes in the graph representation. Ideally, if the assignments are properly 
embedded, the graph should be able to automatically reflect the resulting formula after an assignment. 
Otherwise, adding annotations on nodes to identify live literals and clauses or manually pruning the graph 
(i.e., removing the edges and nodes) are two options.

\paragraph{Solutions}

It seems a perfect self-adapting graph neural network architecture that reflects the assignments
does not exist yet.

So we discuss the second pruning schema that changes the graph structure every time.
Note that there are conceptually two neural networks, $\mathcal{G}$ transforms a formula (graph represented) 
to embeddings, and $\mathcal{P}$ predicts how to choose from the live variables according to the embeddings.
Also note that if conflict-driven clauses learning is not introduced, 
the graph is monotonically shrinking every time.

Consider a complete decision trace that consists of $t$ decisions:
$$ (F_0, G_0, d_0) \to (F_1, G_1, d_1) \cdots \to (F_t, G_t, d_t) $$

where $F_i$ is the formula, $G_i$ is the corresponding graph embedding, and $d_i$ is the decision at time $i$.
A decision $d$ is a pair that consists of a variable and a assigned value (either $\top$ or $\bot$).

If eventually we have SAT, then we assign a reward $1$ for every decision on the this path,
and backprop the two neural networks $\mathcal{G}$ and $\mathcal{P}$ at each level of decisions.

If eventually we have UNSAT or have to backtrack, then we assign reward $-\frac{1}{2^{t-i+1}}$
and backprop the $\mathcal{G}$ and $\mathcal{P}$ at the level $i$.
For example, we assign the most recent decision a 
reward $-\frac{1}{2}$ and backprop the $\mathcal{G}$ and $\mathcal{P}$ at this level; then assign
a reward $-\frac{1}{4}$ and backprop the $\mathcal{G}$ and $\mathcal{P}$ at level $t-1$, and etc.

\clearpage{}

% %% Acknowledgments
% \begin{acks}                            %% acks environment is optional
%                                         %% contents suppressed with 'anonymous'
%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
%   %% acknowledge financial support and will be used by metadata
%   %% extraction tools.
%   This material is based upon work supported by the
%   \grantsponsor{GS100000001}{National Science
%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
%   conclusions or recommendations expressed in this material are those
%   of the author and do not necessarily reflect the views of the
%   National Science Foundation.
% \end{acks}

%% Bibliography
\bibliography{reference}

%% Appendix
% \appendix
% \section{Appendix}

% Text of appendix \ldots

\end{document}